```{r}
library(tidyverse)
library(tidymodels)
library(e1071) #often needed for various statistical tasks
library(ROCR) #for threshold selction
library(glmnet) #for Lasso, ridge, and elastic net models 
library(esquisse)
```

```{r}
library(readr)
parole <- read_csv("parole.csv")
summary(parole)
```

```{r}
parole = parole %>% mutate(male = as_factor(male), race = as_factor(race), state = as_factor(state), crime = as_factor(crime), multiple.offenses = as_factor(multiple.offenses), violator = as_factor(violator))
                           
```

```{r}
parole = parole %>% mutate(male = fct_recode(male, "Female" = "0", "Male" = "1" ))

parole = parole %>% mutate(race = fct_recode(race, "Other Race" = "2", "White" = "1" ))

parole = parole %>% mutate(state = fct_recode(state, "Kentucky" = "2", "Louisiana" = "3", "Virginia" = "4", "Other State" = "1" ))

parole = parole %>% mutate(multiple.offenses = fct_recode(multiple.offenses, "First Offense" = "0", "Multiple Offenses" = "1" ))

parole = parole %>% mutate(crime = fct_recode(crime, "Larceny" = "2", "Drug-Related" = "3", "Driving-Related" = "4", "Other Crime" = "1" ))

parole = parole %>% mutate(violator = fct_recode(violator, "Completed Parole - No Violation" = "0", "Violated Parole" = "1" ))
```

```{r}
summary(parole)
```

```{r}
set.seed(12345)
parole_split = initial_split(parole, prop = 0.70, strata = violator)
train = training(parole_split)
test = testing(parole_split)
```

```{r}
levels(train$violator)
```

The code below reorders levels
```{r}
train = train %>% mutate(violator = fct_relevel(violator, c("Completed Parole - No Violation","Violated Parole")))
levels(train$violator)

```
```{r}
ggplot(parole, aes(x=violator, fill = state)) + geom_bar(position="fill") + theme_bw()
```

```{r}
ggplot(parole, aes(x=violator, y= max.sentence)) + geom_boxplot() + theme_bw()
```

```{r}
parole_model = 
  logistic_reg() %>% #note the use of logistic_reg
  set_engine("glm") #standard logistic regression engine is glm

parole_recipe = recipe(violator ~ state, train) %>%
  step_dummy(all_nominal(), -all_outcomes()) #exclude the response variable from being dummy converted  

logreg_wf = workflow() %>%
  add_recipe(parole_recipe) %>% 
  add_model(parole_model)

parole_fit = fit(logreg_wf, train)
```

```{r}
summary(parole_fit$fit$fit$fit)
```
```{r}
train_model = 
  logistic_reg() %>% #note the use of logistic_reg 
  set_engine("glm") #standard logistic regression engine is glm

train_recipe = recipe(violator ~ state + multiple.offenses + race, train) %>%
  step_dummy(all_nominal(), -all_outcomes())

logreg_wf = workflow() %>%
  add_recipe(train_recipe) %>% 
  add_model(train_model)

train_fit = fit(logreg_wf, train)
```

```{r}
summary(train_fit$fit$fit$fit)
```
```{r}
newdata = data.frame(state = "Louisiana", multiple.offenses = "Multiple Offenses", race = "White")
predict(train_fit, newdata, type="prob")
```
Develop predictions  
```{r}
predictions = predict(train_fit, train, type="prob") #develop predicted probabilities
head(predictions)
```
Let's extract just the "Yes" prediction.  
```{r}
predictions = predict(train_fit, train, type="prob")[2]
head(predictions)
```

Threshold selection  
```{r}
#Change this next line to the names of your predictions and the response variable in the training data frame
ROCRpred = prediction(predictions, train$violator) 

###You shouldn't need to ever change the next two lines:
ROCRperf = performance(ROCRpred, "tpr", "fpr")
plot(ROCRperf, colorize=TRUE, print.cutoffs.at=seq(0,1,by=0.1), text.adj=c(-0.2,1.7))
```
Area under the curve (AUC). AUC is a measure of the strength of the model. Values closer to 1 are better. Can be used to compare models.  
```{r}
as.numeric(performance(ROCRpred, "auc")@y.values)
```

```{r}
#Determine threshold to balance sensitivity and specificity
#DO NOT modify this code
opt.cut = function(perf, pred){
    cut.ind = mapply(FUN=function(x, y, p){
        d = (x - 0)^2 + (y-1)^2
        ind = which(d == min(d))
        c(sensitivity = y[[ind]], specificity = 1-x[[ind]], 
            cutoff = p[[ind]])
    }, perf@x.values, perf@y.values, pred@cutoffs)
}
print(opt.cut(ROCRperf, ROCRpred))
```
Test thresholds to evaluate accuracy  
```{r}
#confusion matrix
#The "No" and "Yes" represent the actual values
#The "FALSE" and "TRUE" represent our predicted values
t1 = table(train$violator,predictions >  0.2015788)
t1
```

Calculate accuracy  
```{r}
(t1[1,1]+t1[2,2])/nrow(train)
```
Can apply trial and error to maximize accuracy (here trying 0.5 as threshold)
```{r}
t1 = table(train$violator,predictions >  0.2)
t1
(t1[1,1]+t1[2,2])/nrow(train)
```

```{r}
t1 = table(train$violator,predictions >  0.3)
t1
(t1[1,1]+t1[2,2])/nrow(train)
```
```{r}
t1 = table(train$violator,predictions >  0.4)
t1
(t1[1,1]+t1[2,2])/nrow(train)
```

```{r}
t1 = table(train$violator,predictions >  0.5)
t1
(t1[1,1]+t1[2,2])/nrow(train)
```


